CrypTool Script \\ Mathematics and Cryptography

### 2002-07-12 script/de/chapters/primes.tex
be remarks

\begin{center}
\fbox{\parbox{15cm}{
    \emph{Albert Einstein\footnotemark:}\\
    Der Fortschritt lebt vom Austausch des Wissens.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  deutscher Physiker und Nobelpreistr"ager, 14.03.1875$-$14.04.1955
}

@@

@@

    dann in der aufsteigenden Folge), so dass sie insgesamt 40 verschiedene Primzahlwerte liefert (dieselben wie die Funktion aus den Punkt 10).

@@

\begin{itemize}
  \item Der deutsche Mathematiker \index{Goldbach Christian} Christian Goldbach (1690-1764) vermutete:
    Jede gerade nat"ur"-liche Zahl gr"o"ser $2$ l"asst sich als die Summe zweier Primzahlen darstellen.
    Mit Computern ist die Goldbachsche Vermutung f"ur alle geraden Zahlen bis $4*10^{14}$ verifiziert\footnote{%
Da"s die Goldbachsche Vermutung wahr ist, d.h. f"ur alle geraden nat"urlichen Zahlen gr"o"ser als $2$ gilt, wird
    heute allgemein nicht mehr angezweifelt. Der Mathematiker J"org Richstein vom Institut f"ur Informatik der 
    Universit"at Gie"sen hat 1999 die geraden Zahlen bis 400 Billionen untersucht und kein Gegenbeispiel gefunden (Siehe 
    \href{http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}
         {\tt http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}).
    \index{Richstein 1999}
    Trotzdem ist das kein allgemeiner Beweis.\\ 
    Da"s die Goldbach’sche Vermutung trotz aller Anstrengungen bis heute nicht bewiesen wurde, f"ordert 
    allerdings einen Verdacht:
    Seit den bahnbrechenden Arbeiten des "osterreichischen Mathematikers Kurt G"odel\index{G\""odel Kurt}
    ist bekannt, dass nicht jeder wahre Satz in der Mathematik auch beweisbar ist (Siehe 
    \href{http://www.mathematik.ch/mathematiker/goedel.html}
         {\tt http://www.mathematik.ch/mathematiker/goedel.html}).
    M"oglicherweise hat Goldbach also Recht,
    und trotzdem wird nie ein Beweis gefunden werden. Das wiederum l"asst sich aber vermutlich auch nicht beweisen.
    },   % footnote
    aber allgemein noch nicht bewiesen\footnote{%
Der englische Verlag {\em Faber} und die amerikanische
    Verlagsgesellschaft {\em Bloomsbury} publizierten 2000 das 1992 erstmal
    ver"offentlichte Buch \glqq Onkel Petros und die Goldbachsche Vermutung\grqq~ von Apostolos Doxiadis (deutsch bei
    L"ubbe 2000 und bei BLT als Taschenbuch 2001). Es ist 
    die Geschichte eines Mathematikprofessors, der daran scheitert, ein mehr als 250 Jahre altes R"atsel zu l"osen.\\
    Um die Verkaufszahlen zu f"ordern, schreiben die beiden  Verlage einen Preis von 1 Million USD aus, wenn 
    jemand die Vermutung beweist -- ver"offentlicht in einer angesehenen mathematischen Fachzeitschrift bis 2004 (Siehe 
    \href{http://www.mscs.dal.ca/~dilcher/Goldbach/index.html}
         {\tt http://www.mscs.dal.ca/\~{}dilcher/Goldbach/index.html}).\\
    Erstaunlicherweise d"urfen nur englische und amerikanische Mathematiker daran teilnehmen.\\
    Der Beweis, der der Goldbach-Vermutung bisher am n"achsten kommt, wurde 1966 von Chen Jing-Run
    bewiesen -- in einer schwer nachvollziehbaren Art und Weise: Jede gerade Zahl gr"o"ser 2 ist die Summe einer 
    Primzahl und des Produkts zweier Primzahlen. Z.B. $20=5+3*5.$\\ 
    Gerade diese Vermutung legt nahe, dass wir auch heute noch nicht in aller Tiefe den Zusammenhang zwischen
    der Addition un der Multiplikation der nat"urlichen Zahlen verstanden haben. \\
    Die wichtigsten Forschungsergebnisse zur Goldbachschen Vermutung sind zusammengefasst in dem von Wang Yuan 
    herausgegebenen Band: \glqq Goldbach Conjecture\grqq, 1984, 
    World Scientific Series in Pure Maths, Vol. 4.
    }.   % footnote
\item Der deutsche \index{Riemann Bernhard} Mathematiker Bernhard Riemann (1826-1866) stellte eine bisher 
    unbewiesene, aber auch nicht widerlegte (?) Formel f"ur die Verteilung von Primzahlen auf, die die Absch"atzung 
    weiter verbessern w"urde.
\end{itemize}

@@

Primzahlzwillinge sind Primzahlen, die den Abstand 2 voneinander haben, zum Beispiel 5 und 7 oder
101 und 103. Primzahldrillinge gibt es dagegen nur eines: 3, 5, 7.  Bei allen anderen Dreierpacks
aufeinanderfolgender ungerader Zahlen ist immer eine durch 3 teilbar und somit keine Primzahl.
\begin{itemize}
  \item Offen ist die Anzahl der Primzahlzwillinge: unendlich viele oder eine begrenzte Anzahl?
  Das bis heute bekannte gr"o"ste Primzahlzwillingspaar ist
  $1.693.965 \cdot 2^{66.443} \pm 1.$
  \item Gibt es eine Formel f"ur die Anzahl der Primzahlzwillinge pro Intervall?
  \item Der Beweis oben zu der Funktion  $f: N \rightarrow Z$ mit $n \mapsto a^{3^n}$  garantiert nur die Existenz einer
    solchen Zahl $a$.  Wie kann diese Zahl $a$ bestimmt werden, und wird sie einen Wert haben, so dass
    die Funktion auch von praktischem Interesse ist?
  \item Gibt es unendlich viele Mersenne-Primzahlen?\index{Primzahlen!Mersenne}\index{Mersenne Marine!Mersenne-Primzahlen}
  \item Gibt es unendlich viele Fermatsche Primzahlen?
  \item Gibt es einen Polynomialzeit-Algorithmus zur Zerlegung einer Zahl in ihre Primfaktoren (vgl.
    \cite[S. 167]{Klee1997})?
    Diese Frage kann man auf die beiden folgenden Fragestellungen aufsplitten:
  \begin{itemize}
        \item Gibt es einen Polynomialzeit-Algorithmus, der entscheidet, ob eine Zahl prim ist?
        \item Gibt es einen Polynomialzeit-Algorithmus, mit dem sich f"ur eine zusammengesetzte Zahl $n$ ein
        nicht-trivialer (d.h. von $1$ und von $n$ verschiedener) Teiler von $n$ berechnen l"asst?
        Vergleiche auch Kapitel \ref{RSABernstein} und Kapitel \ref{NoteFactorisation}.
        \end{itemize}
\end{itemize}
\paragraph{Weitere interessante Themen rund um Primzahlen}
Nicht betrachtet wurden in diesem Kapitel folgende eher
zahlentheoretische Themen wie Teilbarkeitsregeln, Modulo-Rechnung,
modulare Inverse, modulare Potenzen und Wurzeln, chinesischer
Restesatz, Eulersche Phi-Funktion, perfekte Zahlen.

@@

\bibitem[Doxiadis2000]{Dioxadis2000}
    Apostolos Dioxadis, \\
    Onkel Petros und die Goldbachsche Vermutung, Bloomsbury 2000 (deutsch
    bei L"ubbe 2000 und bei BLT als Taschenbuch 2001).

>>>>

\begin{center}
\fbox{\parbox{15cm}{
    \emph{Albert Einstein\footnotemark:}\\
    Der Fortschritt lebt vom Austausch des Wissens.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  deutscher Physiker und Nobelpreistr"ager, 14.03.1879$-$14.04.1955
}

@@

\pagebreak %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    dann in der aufsteigenden Folge), so dass sie insgesamt 40 verschiedene Primzahlwerte liefert (dieselben wie die Funktion aus Punkt 10).

@@

\begin{itemize}
  \item Der deutsche Mathematiker \index{Goldbach Christian} Christian Goldbach (1690-1764) vermutete:
    Jede gerade nat"ur"-liche Zahl gr"o"ser $2$ l"asst sich als die Summe zweier Primzahlen darstellen.
    Mit Computern ist die Goldbachsche Vermutung f"ur alle geraden Zahlen bis $4*10^{14}$ verifiziert\footnote{%
Da"s die Goldbachsche Vermutung wahr ist, d.h. f"ur alle geraden nat"urlichen Zahlen gr"o"ser als $2$ gilt, wird
    heute allgemein nicht mehr angezweifelt. Der Mathematiker J"org Richstein vom Institut f"ur Informatik der 
    Universit"at Gie"sen hat 1999 die geraden Zahlen bis 400 Billionen untersucht und kein Gegenbeispiel gefunden (Siehe 
    \href{http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}
         {\tt http://www.informatik.uni-giessen.de/staff/richstein/de/Goldbach.html}).
    \index{Richstein 1999}
    Trotzdem ist das kein allgemeiner Beweis.\\ 
    Da"s die Goldbach’sche Vermutung trotz aller Anstrengungen bis heute nicht bewiesen wurde, f"ordert 
    allerdings einen Verdacht:
    Seit den bahnbrechenden Arbeiten des "osterreichischen Mathematikers Kurt G"odel\index{G\""odel Kurt}
    ist bekannt, dass nicht jeder wahre Satz in der Mathematik auch beweisbar ist (Siehe 
    \href{http://www.mathematik.ch/mathematiker/goedel.html}
         {\tt http://www.mathematik.ch/mathematiker/goedel.html}).
    M"oglicherweise hat Goldbach also Recht,
    und trotzdem wird nie ein Beweis gefunden werden. Das wiederum l"asst sich aber vermutlich auch nicht beweisen.
    },   % footnote
    aber allgemein noch nicht bewiesen\footnote{%
Der englische Verlag {\em Faber} und die amerikanische
    Verlagsgesellschaft {\em Bloomsbury} publizierten 2000 das 1992 erstmal
    ver"offentlichte Buch \glqq Onkel Petros und die Goldbachsche Vermutung\grqq~ von Apostolos Doxiadis (deutsch bei
    L"ubbe 2000 und bei BLT als Taschenbuch 2001). Es ist 
    die Geschichte eines Mathematikprofessors, der daran scheitert, ein mehr als 250 Jahre altes R"atsel zu l"osen.\\
    Um die Verkaufszahlen zu f"ordern, schreiben die beiden  Verlage einen Preis von 1 Million USD aus, wenn 
    jemand die Vermutung beweist -- ver"offentlicht in einer angesehenen mathematischen Fachzeitschrift bis 2004 (Siehe 
    \href{http://www.mscs.dal.ca/~dilcher/Goldbach/index.html}
         {\tt http://www.mscs.dal.ca/\~{}dilcher/Goldbach/index.html}).\\
    Erstaunlicherweise d"urfen nur englische und amerikanische Mathematiker daran teilnehmen.\\
    Der Beweis, der der Goldbach-Vermutung bisher am n"achsten kommt, wurde 1966 von Chen Jing-Run
    bewiesen -- in einer schwer nachvollziehbaren Art und Weise: Jede gerade Zahl gr"o"ser 2 ist die Summe einer 
    Primzahl und des Produkts zweier Primzahlen. Z.B. $20=5+3*5.$\\ 
    Die wichtigsten Forschungsergebnisse zur Goldbachschen Vermutung sind zusammengefasst in dem von Wang Yuan 
    herausgegebenen Band: \glqq Goldbach Conjecture\grqq, 1984, 
    World Scientific Series in Pure Maths, Vol. 4.
    Gerade diese Vermutung legt nahe, dass wir auch heute noch nicht in aller Tiefe den Zusammenhang zwischen
    der Addition und der Multiplikation der nat"urlichen Zahlen verstanden haben. \\
    }.   % footnote
\item Der deutsche \index{Riemann Bernhard} Mathematiker Bernhard Riemann (1826-1866) stellte eine bisher 
    unbewiesene, aber auch nicht widerlegte (?) Formel f"ur die Verteilung von Primzahlen auf, die die Absch"atzung 
    weiter verbessern w"urde.
\end{itemize}

@@

Primzahlzwillinge sind Primzahlen, die den Abstand 2 voneinander haben, zum Beispiel 5 und 7 oder
101 und 103. Primzahldrillinge gibt es dagegen nur eines: 3, 5, 7.  Bei allen anderen Dreierpacks
aufeinanderfolgender ungerader Zahlen ist immer eine durch 3 teilbar und somit keine Primzahl.
\begin{itemize}
  \item Offen ist die Anzahl der Primzahlzwillinge: unendlich viele oder eine begrenzte Anzahl?
  Das bis heute bekannte gr"o"ste Primzahlzwillingspaar ist
  $1.693.965 \cdot 2^{66.443} \pm 1.$
  \item Gibt es eine Formel f"ur die Anzahl der Primzahlzwillinge pro Intervall?
  \item Der Beweis oben zu der Funktion  $f: N \rightarrow Z$ mit $n \mapsto a^{3^n}$  garantiert nur die Existenz einer
    solchen Zahl $a$.  Wie kann diese Zahl $a$ bestimmt werden, und wird sie einen Wert haben, so dass
    die Funktion auch von praktischem Interesse ist?
  \item Gibt es unendlich viele Mersenne-Primzahlen?\index{Primzahlen!Mersenne}\index{Mersenne Marine!Mersenne-Primzahlen}
  \item Gibt es unendlich viele Fermatsche Primzahlen?
  \item Gibt es einen Polynomialzeit-Algorithmus zur Zerlegung einer Zahl in ihre Primfaktoren (vgl.
    \cite[S. 167]{Klee1997})?
    Diese Frage kann man auf die beiden folgenden Fragestellungen aufsplitten:
  \begin{itemize}
        \item Gibt es einen Polynomialzeit-Algorithmus, der entscheidet, ob eine Zahl prim ist?
        \item Gibt es einen Polynomialzeit-Algorithmus, mit dem sich f"ur eine zusammengesetzte Zahl $n$ ein
        nicht-trivialer (d.h. von $1$ und von $n$ verschiedener) Teiler von $n$ berechnen l"asst?
        \footnote{Vergleiche auch Kapitel \ref{RSABernstein} und Kapitel \ref{NoteFactorisation}.}
        \end{itemize}
\end{itemize}
\paragraph{Weitere interessante Themen rund um Primzahlen}
Nicht betrachtet wurden in diesem Kapitel folgende eher
zahlentheoretische Themen wie Teilbarkeitsregeln, Modulo-Rechnung,
modulare Inverse, modulare Potenzen und Wurzeln, chinesischer
Restesatz, Eulersche Phi-Funktion, perfekte Zahlen.

@@

\bibitem[Doxiadis2000]{Dioxadis2000}
    Apostolos Doxiadis, \\
    Onkel Petros und die Goldbachsche Vermutung, Bloomsbury 2000 (deutsch
    bei L"ubbe 2000 und bei BLT als Taschenbuch 2001).

### 2002-07-12 script/en/chapters/aboutcryptool.tex
misc
be remarks

\parskip 4pt
\vskip + 30 pt
{
In this CrypTool script you will find predominantly mathematically oriented
information for using encryption procedures. The main chapters have been written
by {\em various authors} and are therefore independent from one another. At the end of
most chapters you will find literature and URLs.

@@

The {\em authors} Bernhard Esslinger, Bartol Filipovic, Henrik Koy and Roger Oyono
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Giessen and Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

>>>>

\parskip 4pt
\vskip + 30 pt
{
In this CrypTool script you will find predominantly mathematically oriented
information for using encryption procedures. The main chapters have been written
by {\em various authors} and are therefore independent from one another. At the end of
most chapters you will find literature and web links.

@@

The {\em authors} Bernhard Esslinger, Bartol Filipovic, Henrik Koy,
Roger Oyono and J\"org Cornelius Schneider
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Giessen, Siegen and Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

### 2002-07-12 script/en/chapters/digitalsignatures.tex
be remarks

\newpage
\section{Digital signatures}
\index{Signatures!digital}
(Esslinger B. / Filipovics B. / Schneider J. / Koy H., Update: Juni 2002)

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance, we use a cryptographic
hash function.

@@

All participants who want to \index{X.509}  have X.509 verify that their public
key belongs to a real person consult what is known as a \index{Certification
authority (CA)} certification authority (CA). They prove their identity to this
CA (for example by showing their ID). The CA then issues them an electronic
document (certificate) which essentially contains the names of the certificate-holder
and the CA, the certificate-holder's public key and the validity period
of the certificate. The CA then signs the certificate using its secret key.

>>>>

\newpage
\section{Digital signatures}
\index{Signatures!digital}
(Esslinger B. / Filipovics B. / Schneider J. / Koy H., Update: June 2002)

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use a cryptographic
hash function.

@@

All participants who want to \index{X.509} have a X.509 certificate verifying that their public
key belongs to a real person consult what is known as a
\index{Certification authority (CA)} certification authority (CA). They prove their identity to this
CA (for example by showing their ID). The CA then issues them an electronic
document (certificate) which essentially contains the names of the certificate-holder
and the CA, the certificate-holder's public key and the validity period
of the certificate. The CA then signs the certificate using its secret key.

### 2002-07-12 script/en/chapters/ellipticcurves.tex
be remarks

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / Esslinger B. / Oyono R., April 2000, Update Dec. 2001,
June 2002)

\subsection{Elliptic curce cryptography --- a high-performance substitute for RSA?}

@@

The efficiency of a cryptographic algorithm depends on the key length and
the calculation effort that is necessary to provide a prescribed level of
security. The major advantage of ECC compared to RSA is that it requires
much shorter key lengths. If we assume that the computing power increases
by Moore’s law (i.~e.\ it doubles every 18 months), then the evolution of
the key lengths for secure communication will be as
figure~\ref{ThousandBitMultiplications} (source: Arjen Lenstra und Eric
Verheul:
\href{http://cryptosavvy.com/table.htm}{\texttt{http://cryptosavvy.com/table.htm}}).

% -> Figure 1
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.75]{figures/RSAKeylength}
\caption{Comparison of signing and verification time for RSA and Elliptic Curves} 
\label{RSAKeylength}
\end{center}
\end{figure}

@@

% -> Figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.75]{figures/ThousandBitMultiplications}
\caption{Prognose of the key lengths to be regarded safe for RSA and
  Elliptic Curves\vspace{1ex}} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}

@@

Enough about the role of elliptic curves in pure mathematics. In 1985 Neal
Koblitz and Victor Miller independently suggested using elliptic curves in
cryptography. Elliptic curves have thus also found a concrete practical
application. Another interesting field of application for elliptic curves is for
factorising whole numbers. (For example the RSA cryptography system is based on
the \index{Complexity} difficulty/complexity of finding prime factors of an
extremely large number.) In this area, procedures based on elliptic curves have
been investigated and partially used since 1987 (a study by H.W. Lenstra). There
are also prime number tests based on elliptic curves.

@@

The procedure implements the cryptographic primitives for generating and
verifying signatures for the variations of Nyberg-Rueppel signatures and
\index{DSA} DAS signatures based on elliptic curves (in accordance with draft
no. 8 of the IEEE P1363 work group). This was done in collaboration with the
SECUDE GmbH --- using the above library and the SECUDE SDK.
\index{SECUDE GmbH}

@@

                \href{http://grouper.ieee.org/groups/1363}{\texttt{http://grouper.ieee.org/groups/1363}}
        \item  \hypertarget{Lenstra2}{}
An informative web page to the factorisation with elliptic curves.

@@

There, one finds literature to the topic factorisation with elliptic curves as well as link to other web page.

>>>>

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001,
June 2002)

\subsection{Elliptic curce cryptography --- a high-performance substitute for RSA?}\index{Performance}

@@

The efficiency of a cryptographic algorithm depends on the key length and
the calculation effort that is necessary to provide a prescribed level of
security. The major advantage of ECC compared to RSA is that it requires
much shorter key lengths. If we assume that the computing power increases
by Moore’s law (i.~e.\ it doubles every 18 months), then the evolution of
the key lengths for secure communication will be as
figure~\ref{RSAKeylength} (source: Arjen Lenstra und Eric
Verheul:
\href{http://cryptosavvy.com/table.htm}{\texttt{http://cryptosavvy.com/table.htm}}).

% -> Figure 1
\begin{figure}[h]
\begin{center}
\vspace{1.5cm}\includegraphics[scale=0.75]{figures/RSAKeylength}
\caption{Prognose of the key lengths to be regarded safe for RSA and
  Elliptic Curves\vspace{1ex}} 
\label{RSAKeylength}
\end{center}
\end{figure}

@@

% -> Figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.75]{figures/ThousandBitMultiplications}
\caption{Comparison of signing and verification time for RSA and Elliptic Curves} 
\label{ThousandBitMultiplications}
\end{center}
\end{figure}

@@

Enough about the role of elliptic curves in pure mathematics. In 1985 Neal
Koblitz and Victor Miller independently suggested using elliptic curves in
cryptography. Elliptic curves have thus also found a concrete practical
application. Another interesting field of application for elliptic curves is for
factorising whole numbers. (For example the RSA cryptography system is based on
the \index{Complexity} difficulty/complexity of finding prime factors of an
extremely large number.) In this area, procedures based on elliptic curves have
been investigated and partially used since 1987 (a study by H.W. Lenstra). There
are also prime number tests\index{Prime number!test} based on elliptic curves.

@@

The procedure implements the cryptographic primitives for generating and
verifying signatures for the variations of Nyberg-Rueppel signatures and
\index{DSA} DAS signatures based on elliptic curves (in accordance with draft
no. 8 of the IEEE P1363 work group). This was done in collaboration with the
Secude GmbH --- using the above library and the Secude SDK.
\index{Secude}

@@

                \href{http://grouper.ieee.org/groups/1363}{\texttt{http://grouper.ieee.org/groups/1363}}
        \item  \hypertarget{Lenstra2}{}
An informative web page about factorisation with elliptic curves.

@@

There, one finds literature to the topic factorisation with elliptic curves as well as links to other web page.

### 2002-07-12 script/en/chapters/introduction.tex
be remarks

CrypTool is a program with an extremely comprehensive online help enabling you
to use and analyse cryptographic procedures.\par \vskip + 3pt

For the conventional encryption procedures, automatic analyses are available in CrypTool
that enable you to determine the key with knowledge of the encrypted document
and, possibly, further information (non-encrypted document or language of the
document).\par \vskip + 3pt

CrypTool was developed during the End-User Awareness program in order to
increase employee awareness of IT security and provide them with a deeper
understanding of the term security.\par \vskip + 3pt

A further aim was to enable users to understand the cryptographic procedures
implemented in the Deutsche Bank. In this way, using CrypTool as a reliable
reference implementation of the various encryption procedures (because of using the industry-proven SECUDE Library),
you can test the encryption implemented in other programs. \par \vskip + 3pt

@@

>>>>

CrypTool is a program with an extremely comprehensive online help enabling you
to use and analyse cryptographic procedures within a unified graphical user interface.\par \vskip + 3pt

CrypTool was developed during the end-user awareness program in order to
increase employee awareness of IT security and provide them with a deeper
understanding of the term security.\par \vskip + 3pt

A further aim was to enable users to understand the cryptographic procedures
implemented in the Deutsche Bank. In this way, using CrypTool as a reliable
reference implementation of the various encryption procedures (because of using the industry-proven Secude Library\index{Secude}),
you can test the encryption implemented in other programs. \par \vskip + 3pt

Since then CrypTool has been used for education in companies and
universities and several universities help to develop further project.

@@

Additionally we aimed to describe the newest results, as well as their
origins, about primes and factorization in order to be highly topical.

% Local Variables:
% TeX-master: "../script-en.tex"
% End:

### 2002-07-12 script/en/chapters/moderncryptography.tex
misc
be remarks

A 0-1 knapsack problem can be solved by testing the $ 2^n $   possibilities for 
$ x_1, \dots, x_n $. The best method requires $ O(2^{n/2}) $  operations, which 
for $ n=100 $  with $ 2^{100} \approx 1.27 \cdot 10^{30} $  and  $ 2^{n/2} 
\approx 1.13 \cdot 10^{15} $ represents an insurmountable hurdle for computers.
However, for special $ a_1, \dots, a_n $   the solution is quite easy to find, 
e.g. for $ a_i = 2^{i-1}. $  The binary representation of $ g $ immediately 
delivers $ x_1, \dots, x_n$. In general, the a 0-1 knapsack problem can be 
solved easily if a \index{Permutation} permutation\footnote{A permutation $\pi$ 
of the numbers $1, \dots, n$ is a change in the order in which these numbers are 
listed. For example, a permutation $\pi$ of $(1,2,3)$ is $(3,1,2),$ i.e. 
$\pi(1) = 3$, $\pi(2) =1$ 
and $\pi(3) = 2$.} 
$ \pi $  of $ 1, \dots, n $  exists with $ a_{\pi (j)} > \sum_{i=1}^{j-1} 
a_{\pi(i)} $. If, in addition, $ \pi $ is the identity, i.e. $ \pi(i)=i $ for $ 
i=1,2,\dots,n, $ then the sequence $ a_1, \dots , a_n $ is said to be 
superincreasing.
The following algorithm solves the knapsack problem with a superincreasing 
sequence in the timeframe of $ O(n). $
\begin{center}

@@

At the moment, no algorithm is known that can factorise two prime numbers 
sufficiently quickly for extremely large values (e.g. for several hundred 
decimal places). The quickest algorithms known today
\cite{4Stinson1995}\index{Stinson 1995} 
factorise a compound whole number $ N $ in a time period proportional to $ L(N)= 
e^{\sqrt{\ln (N) \ln (\ln (N))}}. $ 
\vskip +5 pt
\begin{center}
\begin{tabular}{|l||l|l|l|l|l|l|}\hline
$N$ & $ 10^{50} $ & $ 10^{100} $ & $ 10^{150} $ & $ 10^{200} $ & $ 10^{250} $ & 
$ 10^{300} $ \\ \hline
$L(N)$ & $ 1.42 \cdot 10^{10} $ &  $ 2.34  \cdot 10^{15} $ &  $ 3.26 \cdot 
10^{19} $ &  $ 1.20 \cdot 10^{23} $ &  $ 1.86 \cdot 10^{26} $ &  $ 1.53 \cdot 
10^{29} $ \\ \hline
\end{tabular}
\end{center}
\vskip +5 pt
As long as no better algorithm is found, this means that values of the order of 
magnitude $ 100 $ to $ 200 $ decimal places are currently safe. Estimates of the 
current computer technology indicate that a number with $100$ decimal places 
could be factorised in approximately two weeks at justifiable cost. Using an 
expensive configuration (e.g. of around 10 million US dollars), a number with 
$150$ decimal places could be factorised in about a year. A $200-$digit number 
should remain impossible to factorise for a long time to come, unless there is a 
mathematical breakthrough. For example, it would take about 1000 years to 
decompose a 200-digit number into prime factors using the existing algorithms; 
this applies even if $ 10^{12} $  operations can be performed per second, which 
is beyond the performance of current technology and would cost billions of 
dollars in development costs. However, you can never be sure that there won't be 
a mathematical breakthrough tomorrow.

@@

\index{Key agreement!Diffie-Hellman} \index{Diffie Whitfield} The mechanisms and 
algorithms of classical cryptography only take effect when the subscribers have 
already exchanged the secret key. In classical cryptography, you cannot avoid 
exchanging secrets without encrypting them. Transmission safety here must be 
achieved using non-cryptographic methods. We say that we need a secret channel 
for exchanging secrets. This channel can be realised either physically or 
organisationally. \\
What is revolutionary about modern cryptography is, amongst other things, that 
you no longer need secret channels: You can agree secret keys using non-secret, 
i.e. public channels. \\
One protocol that solves this problems is that of Diffie and Hellman \\

@@

\vskip +10 pt
{\bf Diffie Hellman} (Key agreement).
\vskip +20 pt

@@

           \bibitem[Shamir1982]{Shamir1982} Shamir A.:
           \emph{A polynomial time algorithm for breaking the basic Merkle-
Hellman Cryptosystem.}\\
           Symposium on Foundations of Computer Science (1982), 145-152.
           \index{Shamir Adi}
%already defined in elementaryNumberTheory.inc
       \bibitem[Stinson1995]{4Stinson1995} Stinson D.R.:
           \emph{Cryptography.}  \index{Stinson 1995} \\
           CRC Press, Boca Raton, London, Tokyo, 1995.\index{Stinson 1995}
\end{thebibliography}
\section*{URLs} \addcontentsline{toc}{subsection}{URLs}
\begin{enumerate}
   \item 
\href{http://www.geocities.com/st_busygin/clipat.html}{http://www.geocities.com/st\_busygin/clipat.html }
\end{enumerate}

>>>>

A 0-1 knapsack problem can be solved by testing the $ 2^n $   possibilities for 
$ x_1, \dots, x_n $. The best method requires $ O(2^{n/2}) $  operations, which 
for $ n=100 $  with $ 2^{100} \approx 1.27 \cdot 10^{30} $  and  $ 2^{n/2} 
\approx 1.13 \cdot 10^{15} $ represents an insurmountable hurdle for computers.
However, for special $ a_1, \dots, a_n $   the solution is quite easy to find, 
e.g. for $ a_i = 2^{i-1}. $  The binary representation of $ g $ immediately 
delivers $ x_1, \dots, x_n$. In general, the a 0-1 knapsack problem can be 
solved easily if a \index{Permutation} permutation\footnote{A permutation\index{Permutation} $\pi$ 
of the numbers $1, \dots, n$ is a change in the order in which these numbers are 
listed. For example, a permutation $\pi$ of $(1,2,3)$ is $(3,1,2),$ i.e. 
$\pi(1) = 3$, $\pi(2) =1$ 
and $\pi(3) = 2$.} 
$ \pi $  of $ 1, \dots, n $  exists with $ a_{\pi (j)} > \sum_{i=1}^{j-1} 
a_{\pi(i)} $. If, in addition, $ \pi $ is the identity, i.e. $ \pi(i)=i $ for $ 
i=1,2,\dots,n, $ then the sequence $ a_1, \dots , a_n $ is said to be 
superincreasing.
The following algorithm solves the knapsack problem with a superincreasing 
sequence in the timeframe of $ O(n). $
\begin{center}

@@

At the moment, no algorithm is known that can factorise two prime numbers 
sufficiently quickly for extremely large values (e.g. for several hundred 
decimal places). The quickest algorithms known today
\cite{4Stinson1995}\index{Stinson 1995} 
factorise a compound whole number $ N $ in a time period proportional to $ L(N)= 
e^{\sqrt{\ln (N) \ln (\ln (N))}}. $ 
\vskip +5 pt
\begin{center}
\begin{tabular}{|l||l|l|l|l|l|l|}\hline
$N$ & $ 10^{50} $ & $ 10^{100} $ & $ 10^{150} $ & $ 10^{200} $ & $ 10^{250} $ & 
$ 10^{300} $ \\ \hline
$L(N)$ & $ 1.42 \cdot 10^{10} $ &  $ 2.34  \cdot 10^{15} $ &  $ 3.26 \cdot 
10^{19} $ &  $ 1.20 \cdot 10^{23} $ &  $ 1.86 \cdot 10^{26} $ &  $ 1.53 \cdot 
10^{29} $ \\ \hline
\end{tabular}
\end{center}
\vskip +5 pt
As long as no better algorithm is found, this means that values of the order of 
magnitude $ 100 $ to $ 200 $ decimal places are currently safe. Estimates of the 
current computer technology indicate that a number with $100$ decimal places 
could be factorised in approximately two weeks at justifiable cost. Using an 
expensive configuration (e.g. of around 10 million US dollars), a number with 
$150$ decimal places could be factorised in about a year. A $200-$digit number 
should remain impossible to factorise for a long time to come, unless there is a 
mathematical breakthrough. For example, it would take about 1000 years to 
decompose a 200-digit number into prime factors using the existing algorithms; 
this applies even if $ 10^{12} $  operations can be performed per second, which 
is beyond the performance\index{Performance} of current technology and would cost billions of 
dollars in development costs. However, you can never be sure that there won't be 
a mathematical breakthrough tomorrow.

@@

\index{Key agreement (key exchange)!Diffie-Hellman} \index{Diffie Whitfield} The mechanisms and 
algorithms of classical cryptography only take effect when the subscribers have 
already exchanged the secret key. In classical cryptography, you cannot avoid 
exchanging secrets without encrypting them. Transmission safety here must be 
achieved using non-cryptographic methods. We say that we need a secret channel 
for exchanging secrets. This channel can be realised either physically or 
organisationally. \\
What is revolutionary about modern cryptography is, amongst other things, that 
you no longer need secret channels: You can agree secret keys using non-secret, 
i.e. public channels. \\
One protocol that solves this problems is that of Diffie and Hellman \\

@@

\vskip +10 pt
{\bf Diffie Hellman key agreement.}
\vskip +20 pt

@@

           \bibitem[Shamir1982]{Shamir1982} Shamir A.:
           \emph{A polynomial time algorithm for breaking the basic Merkle-
Hellman Cryptosystem.}\\
           Symposium on Foundations of Computer Science (1982), 145-152.
           \index{Shamir Adi}
%already defined in elementaryNumberTheory.inc
       \bibitem[Stinson1995]{4Stinson1995} Stinson D.R.:
           \emph{Cryptography.}  \index{Stinson 1995} \\
           CRC Press, Boca Raton, London, Tokyo, 1995.\index{Stinson 1995}
\end{thebibliography}
\section*{Web links} \addcontentsline{toc}{subsection}{Web links}
\begin{enumerate}
   \item 
\href{http://www.geocities.com/st_busygin/clipat.html}{http://www.geocities.com/st\_busygin/clipat.html }
\end{enumerate}

% Local Variables:
% TeX-master: "../script-en.tex"
% End:
